\textbf{Species and genus plant classification with NEON hyperspectral remote sensing data}

### 2020-11-02 R/clean_spectral_data.Rmd
histogram figure spectra per species
need to rearrange the figures though in the text so study area comes first

```{r, fig.width=12}
# plot interpolated spectra to see their starting wavelengths
interp_wl %>% 
  dplyr::filter(spectraID %in% c("NEON.PLA.D07.MLBS.00412_2018", 
                                 "NEON.PLA.D01.HARV.04350_2018",
                                 "NEON.PLA.D08.LENO.08750_2015",
                                 "NEON.PLA.D01.BART.06298_2019")) %>% 
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra w interpolated wavelengths: Zoom in on beginning of spectra",
       x = "wavelength [nm]",
       color = "spectrum ID") + 
  # zoom in on beginning of spectra
  lims(x = c(380,400), y = c(0,0.1)) +
  theme_bw()

@@

# plot histogram with count per genusSpecies. color bars based on whether
# # samples exceeds the threshold
ggplot(taxonCounts, aes(y = reorder(genusSpecies, n), x = n)) + 
                          #reorder(taxonID, n), x = n)) + 
  geom_bar(stat = "identity", aes(fill = countGT)) + 
  labs(title = paste("Histogram of samples per species.",
                     speciesCountTotal, "total species;",
                     speciesCountGT, "species with greater than",
                     thresh, "samples"),
       y = "Scientific Name (Genus + Species)",
       fill = paste("Count >", thresh)) + 
  theme_bw() + 
  scale_fill_manual(values = c("#d95f0e", "#7bccc4"))

@@

>>>>

```{r, fig.width=12}
# plot interpolated spectra to see their starting wavelengths
interp_wl %>% 
  dplyr::filter(spectraID %in% c("NEON.PLA.D11.CLBJ.00496_2018", 
                                 "NEON.PLA.D01.HARV.09041_2018",
                                 "NEON.PLA.D08.DELA.03903_2017")) %>% 
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra w interpolated wavelengths: Zoom in on beginning of spectra",
       x = "wavelength [nm]",
       color = "spectrum ID") + 
  # zoom in on beginning of spectra
  lims(x = c(380,400), y = c(0,0.1)) +
  theme_bw()

@@

# plot histogram with count per genusSpecies. color bars based on whether
# # samples exceeds the threshold
ggplot(taxonCounts, aes(y = reorder(genusSpecies, n), x = n)) + 
                          #reorder(taxonID, n), x = n)) + 
  geom_bar(stat = "identity", aes(fill = countGT)) + 
  labs(title = paste("Histogram of samples per species.",
                     speciesCountTotal, "total species;",
                     speciesCountGT, "species with greater than",
                     thresh, "samples"),
       y = "Scientific Name (Genus + Species)",
       fill = paste("Count >", thresh)) + 
  theme_bw() + 
  scale_fill_manual(values = c("#d95f0e", "#7bccc4"))
```

```{r}
# Histogram with just the taxons that have enough samples 
samplesCountGT <- taxonCounts %>% 
  filter(countGT == TRUE)

# plot histogram with count per genusSpecies. color bars based on whether
# # samples exceeds the threshold
ggplot(samplesCountGT, aes(y = reorder(genusSpecies, n), x = n)) + 
  geom_bar(stat = "identity", fill = "#7E9E8F") + 
  labs(y = "Scientific name\n",
       x = "\nNumber of spectra") + 
  #theme(axis.text = element_text(size = 12)) +
  theme_bw(base_size=18)

@@

```{r}
# table with species info and counts

library(taxize)

taxon_info <- all_spectra %>% 
  select(genusSpecies, taxonID) %>% 
  distinct()

taxons_GT <- taxonCounts %>% 
  # keep taxons that have enough samples 
  filter(countGT == TRUE) %>% 
  # remove the column that indicates if the taxon has enough samples
  select(-countGT) %>% 
  # merge in the taxonID information 
  left_join(taxon_info, by="genusSpecies")

# for(i in 1:nrow(taxons_GT)){
#   print(i)
#   taxize::sci2comm(sci = taxons_GT[i]$genusSpecies)
# }

```

### 2020-11-02 analysis/paper/paper.Rmd
histogram figure spectra per species
need to rearrange the figures though in the text so study area comes first

Each series of wavelengths increments by 5nm from each starting wavelength to yield 426 spectral bands. 
Most of them started at 384nm, so we linearly interpolated each spectrum to have the same 426 bands starting at 384nm and ending at 2512nm with 5nm increments.
We filtered out any spectra with a starting wavelength of 347nm, which were all collected at the Pu'u Maka'ala Natural Area Reserve (PUUM) NEON site, Pacific Tropical Domain 20, Hawaii, since the interpolation would require too much unavailable data outside the available range.
We flagged "bad bands" from 1340-1445nm, 1790-1995nm, and greater than 2400nm so they could be excluded from later analysis.
These wavelength ranges feature noisy reflectance data due to high water absorption [@bajorski2011].

Next, we removed samples belonging to species classes with few samples. 
We figured that a small sample size would lead to a poor representation in model training. 
From the initial collection of 146 species, 50 species had at least 20 samples. 
We randomly assigned each spectrum to the training or validation sets, using an 80/20 split. All 50 species (and therefore, all genus values) were represented in each set.

@@

>>>>

Each series of wavelengths increments by 5nm from each starting wavelength to yield 426 spectral bands. 
Most of them started at 384nm, so we linearly interpolated each spectrum to have the same 426 bands starting at 384nm and ending at 2512nm with 5nm increments.
Spectra from the Pu'u Maka'ala Natural Area Reserve (PUUM) NEON site in Hawaii had a starting wavelength of 347nm. 
This is because NEON sub-contracted the PUUM flights to the Global Airborne Observatory, which has a spectrometer with a different wavelength range compared to the NIS payloads *([cite?](https://www.neonscience.org/observatory/observatory-blog/remote-sensing-partnership-made-paradise))*. 
We filtered out any spectra with a starting wavelength of 347nm, since the interpolation would require too much unavailable data outside the available range.
We flagged "bad bands" from 1340-1445nm, 1790-1995nm, and greater than 2400nm so they could be excluded from later analysis.
These wavelength ranges feature noisy reflectance data due to high water absorption [@bajorski2011].

Next, we removed samples belonging to species classes with few samples. 
We figured that a small sample size would lead to a poor representation in model training. 
From the initial collection of 146 species, 50 species had at least 20 samples (Figure \@ref(fig:fig-histSpecies)). 
We randomly assigned each spectrum to the training or validation sets, using an 80/20 split. All 50 species (and therefore, all genus values) were represented in each set.

@@

```{r fig-histSpecies, fig.cap ="Histogram: spectra count per species", echo=FALSE,  out.width = "450px", message = FALSE, warning=FALSE}
library(here)
knitr::include_graphics(here::here("analysis","figures", "hist_samples_per_species.pdf"))
```
